### 2023-03-13 README.Rmd
Add README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# antwincomm

<!-- badges: start -->

<!-- badges: end -->

Characterizing the marine winter predator community (seabirds and marine mammals) around the northern Antarctic Peninsula (including the Bransfield Strait and South Shetland Islands) using observations collected during the 2012-2016 NOAA/NSF austral winter cruises on the RV Ice Breaker Nathaniel B. Palmer.

*Work in progress!!*

## Code

All code under development for **antwincomm** can be found in this repository. The analysis pipeline [uses `targets`](https://books.ropensci.org/targets/) to manage workflow caching and dependencies. Better documentation to follow (promise). This repository will be archived on Zenodo or somewhere similar at publication.

## Data

The **antwincomm** data are too large for GitHub, so they're stored in the [antwincomm Open Science Framework project](https://osf.io/hwnvy/). Currently private, will be made public no later than time of publication. For access to the OSF project, please contact Max Czapanskiy (max [dot] czapanskiy [at] noaa [dot] gov).

For the code to work, you must download data/WAMLR from the OSF project into data/WAMLR in your local copy of the repository. You can do that manually or by running the following lines of code. Until the project is made public, though, you'll first have to [authenticate `osfr`](https://docs.ropensci.org/osfr/articles/auth.html).

```{r eval=FALSE}
dir.create(here::here("data"))
osf_wamlr <- osfr::osf_retrieve_file("6407bf6228e5c503cc9376b3")
osfr::osf_download(osf_wamlr, path = here::here("data"), recurse = TRUE)
```

## Outputs

See the [project website](https://flukeandfeather.github/io/antwincomm) for a list of reports documenting the analysis pipeline.