### 2023-03-26 reports/01_predators.qmd
Clean up ice and reports

---
title: "Predator summary"
format:
  html:
    code-fold: true
editor: visual
output-dir: "docs"
---

```{r}
#| label: setup
#| include: false

library(sf)
library(targets)
library(tidyverse)

tar_load(c("predators_agg", "ant_sf", "effort_sf"), 
         store = here::here("_targets"))
source(here::here("R", "sp.R"))
```

## Antarctic Peninsula winter predators

### Counts

The five most abundant species.

```{r}

top_five <- predators_agg %>% 
  group_by(species) %>% 
  summarize(total_count = sum(count)) %>% 
  arrange(desc(total_count)) %>%
  slice(1:5)

top_five
```

```{r}

top_five_plot_data <- predators_agg %>% 
  semi_join(top_five, by = "species") %>% 
  group_by(year, species) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  mutate(species = fct_reorder(species, total, .fun = sum),
         year = factor(year, levels = 2016:2012))
year_labels <- top_five_plot_data %>% 
  filter(species == "SNPT") %>% 
  arrange(desc(year)) %>%                   # Make the labels show up at
  mutate(total = cumsum(total) - total / 2) # correct y-axis position

ggplot(top_five_plot_data, aes(species, total, group = year)) +
  geom_col(color = "black", fill = "grey80") +
  geom_text(aes(label = year), year_labels) +
  labs(y = "Total species count") +
  theme_classic() +
  theme(axis.title.x = element_blank())
```

### Maps

Survey effort. 21 out of the 19404 intervals were recorded as nmi \> 1000 (as opposed to \~1 nmi for the rest). Those intervals excluded from the survey effort.

```{r}
#| label: survey_effort
#| warning: false

effort_stars <- effort_sf %>% 
  filter(nmi < 1000) %>% 
  rasterize_counts(count_col = "nmi",
                   res = 2.5e4)
map_lim <- st_bbox(c(xmin = -65, xmax = -55, ymin = -65, ymax = -58),
                       crs = "EPSG:4326") %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.2)
ant_basemap(map_lim) + 
  stars::geom_stars(aes(fill = nmi), data = effort_stars) +
  scale_fill_distiller(palette = "PuRd", 
                       direction = 1, 
                       na.value = NA) +
  scale_x_continuous(breaks = seq(-65, -55, by = 5)) +
  scale_y_continuous(breaks = seq(-65, -58, by = 2)) +
  coord_ant(map_lim) +
  facet_wrap(vars(year)) +
  expand_limits(fill = 0)

```

Distribution maps of the five most abundant species in each survey year, aggregated to 25km bins.

```{r}
#| label: topfive_map
#| warning: false
#| column: screen
#| out-width: 100%
#| fig-width: 6
#| fig-dpi: 600

topfive_sf <- predators_agg %>% 
  semi_join(top_five, by = "species") %>% 
  latlon_to_sf(coords = c("lon_mean", "lat_mean"))

topfive_stars <- rasterize_counts(topfive_sf, res = 2.5e4)

map_lim <- st_bbox(c(xmin = -65, xmax = -55, ymin = -66, ymax = -60),
                       crs = "EPSG:4326") %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.1)

ant_basemap(map_lim) +
  stars::geom_stars(aes(fill = log(count + 1)), data = topfive_stars) +
  scale_fill_distiller(palette = "PuRd",
                       direction = 1,
                       na.value = NA) +
  scale_x_continuous(breaks = c(-65, -60, -55)) +
  scale_y_continuous(breaks = c(-66, -64, -62, -60)) +
  coord_ant(map_lim) +
  facet_grid(species ~ year) +
  expand_limits(fill = 0) +
  theme(axis.text.x = element_text(size = 7))
```

Distribution maps of five species of interest in each survey year, also aggregated to 25km bins.

```{r}
#| label: soi_map
#| warning: false
#| column: screen
#| out-width: 100%
#| fig-width: 6
#| fig-dpi: 600

soi <- c("GEPN", "BLPT", "ANTE", "LESE", "KIWH")

soi_sf <- predators_agg %>% 
  filter(species %in% soi) %>% 
  latlon_to_sf(coords = c("lon_mean", "lat_mean"))

soi_stars <- rasterize_counts(soi_sf, res = 2.5e4)

map_lim <- st_bbox(soi_sf) %>% 
  project_bbox() %>% 
  expand_bbox(factor = 1.2)

ant_basemap(map_lim) +
  stars::geom_stars(aes(fill = log(count + 1)), data = soi_stars) +
  scale_fill_distiller(palette = "PuRd",
                       direction = 1,
                       na.value = NA) +
  scale_x_continuous(breaks = c(-65, -60, -55)) +
  scale_y_continuous(breaks = c(-64, -62, -60, -58)) +
  facet_grid(species ~ year) +
  expand_limits(fill = 0) +
  coord_ant(map_lim) +
  theme(axis.text.x = element_text(size = 7))
```