### 2024-05-30 analysis/paper/paper.qmd
paper: methods

---
title: "Sea-ice and macrozooplankton distribution as determinants of top predator community structure in Antarctic winter"
author:
  - Max F. Czapanskiy:
      orcid: 0000-0002-6302-905X
      institute:
        - ims
        - fed
        - esd1
  - Jarrod A. Santora:
      correspondence: "yes"
      email: jarrod.santora@noaa.gov
      institute: 
        - fed
        - dam
  - Kimberly S. Dietrich:
      institute: 
        - ims
        - fed
  - Megan A. Cimino:
      orcid: 0000-0002-1715-2903
      institute:
        - ims
        - esd1
  - Elliott L. Hazen:
      institute: esd1
  - Christian S. Reiss:
      institute: esd2
  - Richard R. Veit:
      institute: cuny
institute:
  - ims:
      name: Institute of Marine Sciences, University of California Santa Cruz
      address: Santa Cruz, California 95060, USA
  - fed: 
      name: Fisheries Ecology Division, Southwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration
      address: Santa Cruz, California 95060, USA
  - esd1:
      name: Ecosystem Science Division, Southwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration
      address: Monterey, California 93940, USA
  - dam:
      name: Department of Applied Math, University of California Santa Cruz
      address: Santa Cruz, California 95060, USA
  - esd2:
      name: Ecosystem Science Division, Southwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration
      address: La Jolla, California 92037, USA
  - cuny:
      name: City University of New York, College of Staten Island, Biology Department
      address: Staten Island, New York 10314, USA
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/marine-ecology-progress-series.csl" # Insert path for the bib-style
abstract: |
  The Antarctic Peninsula marine ecosystem is highly productive, with large 
  populations of commercially and ecologically important species including 
  Antarctic krill (*Euphausia superba*), Adelié Adélie penguins (*Pygoscelis 
  adeliae*), and crabeater seals (*Lobodon carcinophagus*). The ecology of the 
  peninsula is rapidly changing due to accelerating climate change and fishing 
  pressure. Systematic ecosystem surveys have focused on austral spring and 
  summer, leaving an information gap on winter ecosystem dynamics. Using data 
  from 5 consecutive ecosystem surveys, we quantified the composition and 
  distribution of winter predator communities and investigated the physical and 
  biological influences on community structure. Seabirds and marine mammals 
  clustered into 3 communities: an ice-associated community represented by 
  Adélie penguins and crabeater seals; a diverse marginal ice zone community
  dominated by fur seals and several species of seabirds including 3 petrels, 
  kelp gulls (*Larus dominicanus*), and Antarctic terns (*Sterna vittata*); and 
  an open water community consisting of Southern fulmars (*Fulmarus 
  glacialoides*) and 4 species of petrels. These communities were distributed 
  along an environmental gradient ranging from ice-covered, cold, saline waters 
  to ice-free, warmer, and fresher water with greater chlorophyll 
  concentrations. Predator communities were also associated with different 
  communities of macrozooplankton: ice-associated predators with an extremely 
  diverse assemblage of typically mesopelagic zooplankton; marginal ice zone 
  predators with a community of large-bodied euphausiids community (*E. superba*
  , *E. crystallorophias*); and open water predators with a community of small-
  bodied euphausiids (*Thysanoessa macrura*). Our synthesis of integrated winter 
  predator and macrozooplankton communities relative to sea-ice concentration 
  provides reference points for future ecosystem assessments within this rapidly 
  changing region.
keywords: |
  Biogeography; Krill; Marine mammal; Polar regions; Sea ice; Seabird; 
  Zooplankton
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

<!-- The actual document text starts here: -->

# Introduction

The marine ecosystem surrounding the Antarctic Peninsula is highly productive and rapidly changing [@clarkeClimateChangeMarine2007; @montes-hugoRecentChangesPhytoplankton2009]. Integrated assessments and syntheses across trophic levels will help us better understand variability of the ecosystem and potential impacts of climate change, such as effects of phenological shifts and carry-over effects between seasons. Variability in the spring biomass and distribution of Antarctic phytoplankton, abundance and diversity of zooplankton, and population dynamics of upper trophic level species (e.g. reproductive success and recruitment) are controlled, in part, by physical and biogeochemical processes during the preceding winter [@hinkeDailyActivityMinimum2011; @sabaWinterSpringControls2014; @meyerWinterPackiceZone2017]. However, systematic ecosystem surveys have focused on austral spring and summer due to more favorable weather. Thus, there is a paucity of information on winter ecosystem dynamics (i.e. diversity and distribution of communities). Since climate-related effects will likely accelerate and percolate through the ecosystem, it is important to document what these effects are, including anthropogenic stressors.

Winter-time climate change near the northern Antarctic Peninsula and fisheries pressure are increasing faster than in other seasons. Atmospheric temperatures around the Antarctic Peninsula increased faster than anywhere else in the southern hemisphere over the last 50 yr, with the most rapid warming occurring during austral winter [@jonesSixtyYearsWidespread2019; @ducklowMarinePelagicEcosystem2022]. Warmer winters, combined with stronger winds, have reduced the duration of sea-ice coverage in the region by an average of \>90 d yr^–1^ [@ducklow2013west], leading to reduced biodiversity and biomass at the base of the food web, including phytoplankton, microzooplankton, and larval fish [@linDeclinePlanktonDiversity2021; @corsoClimateDrivesLongterm2022]. Further, recent record low winter sea-ice coverage in the Antarctic Peninsula suggests that the rate of warming and overall resilience of the marine ecosystem may have been underestimated [@purichRecordLowAntarctic2023]. These winter ecosystem changes have negatively impacted predator populations, including those of penguins [@trivelpiece2011variability] and fur seals [@krauseRapidPopulationCollapse2022; @forcadaNinetyYearsChange2023] that had been recovering. Concurrently, the annual commercial catch of Antarctic krill Euphausia superba increased 4-fold from 2010 to 2020 [@kawaguchi2020krill]. As winter sea-ice extent and concentration decrease, the winter fishery activities have expanded farther south and have become more spatially concentrated [@nicol2016fishery]. This spatial shift has resulted in greater overlap between fisheries and predators at local scales [@hinkeIdentifyingRiskConcurrent2017]. The increasing overlap between commercial fisheries, and foraging seals and penguins, brings into question whether negative impact on predators will increase or decrease [@wattersLongtermObservationsAntarctica2020]. Effective ecosystem-based management of the Antarctic Peninsula is thus hindered by limited data on winter-time community structure and foraging interactions among species [@bestleyMarineEcosystemAssessment2020].

The winter communities of seabirds and marine mammals in the northern Antarctic Peninsula is a mix of overwintering and seasonal members [@santoraEnvironmentalDeterminantsTop2014]. These species include a variety of ice-obligate, ice-avoidant, and intermediate species. Ice-obligate (pagophilic) species consist primarily of swimmers, such as crabeater seals *Lobodon carcinophaga*, Adélie penguins *Pygoscelis adeliae*, and leopard seals *Hydrurga leptonyx*. Most of these either follow the ice edge north as it advances in winter [@thiebotAdeliePenguinsExtensive2019] or remain in the pack ice [@burnsWinterHabitatUse2004] all year. Pagophilic predators exploit prey resources associated with pack ice and rely on ice floes to haul out for rest, avoid predators, and reproduce [@southwellTimingPuppingPackice2003; @ainleyBirdsAntarcticSea2017; @besterAntarcticMarineMammals2017]. The ice-avoidant (open water) species include several fulmarine petrels, such as southern fulmars *Fulmarus glacialoides*, cape petrels *Daption capense*, and blue petrels *Halobaena caerulea*. This guild is abundant in the region during the austral summer [@santoraSpatiotemporalPersistenceTop2013; @warwick-evansMultiscaleAssessmentDistribution2021], but many individuals migrate farther north in winter [@delordSpeciesSpecificForaging2016]. These seabirds have broader wing morphology adapted for efficient flight in strong, consistent winds [@ainleySpeciesHabitatRelationshipsAntarctic1993]. In contrast, intermediate between these 2 guilds, snow petrels *Pagodroma nivea* and Antarctic petrels *Thalassoica antarctica* have narrower wings and forage in areas of marginal ice coverage [@santoraEnvironmentalDeterminantsTop2014]. During winter, they are largely confined to the marginal ice zone [@delordAntarcticPetrelsIce2020; @delordSpeciesSpecificForaging2016].

At fine spatial scales (\<10 km), the structure of the top predator community of the Antarctic Peninsula in winter is driven by both physical (sea ice coverage) and biological (prey abundance and distribution) factors [@fraserIceEdgesSeabird1986; @reissOverwinterHabitatSelection2017]. The mesoscale (10s to 100s km) distribution of macrozooplankton communities is more geographically consistent in winter than summer, independent of ice conditions [@dietrich2021winter], whereas the fine-scale distribution of seabird communities dynamically responds to ice coverage in all seasons [@ainleySpeciesHabitatRelationshipsAntarctic1993]. Five winter communities of macrozooplankton have previously been identified in this region [@dietrich2021winter]. Two of these communities contain smaller euphausiids (*Thysanoessa macrura*, *E. frigida*, *E. triacantha*) and amphipods (*Primno*, *Themisto*) and occupy the region offshore of the shelf, north and west of the South Shetland Islands. Within the Bransfield Strait, a community dominated by large euphausiids (*E. superba*, *E. crystallorophias*) occurs in the western part of the strait, while a highly diverse assemblage including chaetognaths, *Calanoides acutus*, and siphonophores occurs in the eastern strait. A fifth macrozooplankton community is ubiquitous within this region and contains pelagic tunicates (*Salpa thompsoni*) and gastropods (*Clione limacina*). If the distribution of prey communities influences structure of predator communities, then we would expect associations between the open water seabird guild and the off-shelf macrozooplankton communities (called 2a and 2b by @dietrich2021winter), as well as the pagophilic penguin/pinniped predator community with the Bransfield Strait macrozooplankton communities (called 3a and 3b by @dietrich2021winter). Conversely, if the physical factors influencing predator community distribution at the fine-scale also drive mesoscale patterns, then predator communities would be associated directly with ice coverage (and perhaps hydrology) rather than macrozooplankton communities.

Here, we investigated the ecological determinants of top predator community structure in the Antarctic Peninsula during winter (August to September) using data from 5 consecutive ecosystem surveys in 2012–2016. The study period coincided with several high and low sea-ice years, and captured natural variability in the extent and distribution of the winter marginal ice zone habitat within the northern Antarctic Peninsula marine ecosystem [@santoraEnvironmentalDeterminantsTop2014; @reissOverwinterHabitatSelection2017; @dietrich2021winter]. Antarctic sea ice reached a record high in 2014, before the current regime of rapid declines [@parkinson40yRecordReveals2019], and an extreme El Niño event in 2015/2016 had substantial impacts on phytoplankton and zooplankton communities [@costaLargeDiatomBloom2021; @walshExtremeNinoSouthern2023]. We clustered top predator observations into distinct communities following the methods used for macrozooplankton communities from the same surveys [@dietrich2021winter]. We quantified associations between the top predator communities and (1) physical environmental factors and (2) macrozooplankton communities to test the hypothesis that community structure at lower trophic levels, combined with sea-ice extent, determines community structure at higher trophic levels. Our winter ecosystem observations have potential implications for future climate and fishery impact assessments that emphasize relationships between the physical environment, macrozooplankton distribution, and top predator community structure.

# Materials and Methods

## Survey area data

### Survey design

We surveyed the South Shetland Island and Elephant Island regions during the austral winters (August to September) of 2012–2016 as part of the US Antarctic Marine Living Resources (AMLR) program aboard the RV Ice Breaker (RVIB) ‘Nathaniel B. Palmer’ (Fig. 1). The main objectives of the survey were to map hydrographic conditions, study the distribution and abundance of fish, krill, and other zooplankton through acoustic methods and net hauls, and describe the spatial distribution of seabirds and marine mammals. We followed methods consistent with the long-term austral summer surveys conducted by the US AMLR program [@reissOverwinterHabitatSelection2017]. We sampled between 33 and 108 stations per year, dependent on ice conditions. Stations were spaced approximately 25–30 km apart, and the full survey area covered an area of approximately 120000 km^2^.

### Visual surveys

Sighting surveys for seabirds and marine mammals were conducted from the bridge of the ship, following established methodologies described in previous studies [@santoraEnvironmentalDeterminantsTop2014]. Two observers collected observations simultaneously during daylight hours. One observer recorded seabirds, while the other scanned for marine mammals in the water and on the ice using binoculars. Observers used 10 × 50 binoculars and image-stabilized 20 × 60 binoculars. The pilot house/bridge of the RVIB ‘Nathaniel B. Palmer’ is 54 ft (\~16.5 m) above sea level. Observations only stopped when the ship could not move forward though the sea ice, and no observations were made during station sampling.

During each survey, we continuously monitored and classified sea-ice conditions according to the international sea ice protocol (ASPeCt) [@worbyPartObservationalTechnique1999; @santoraSpatiotemporalPersistenceTop2013; @reissOverwinterHabitatSelection2017]. This involved estimating the percentage of ice cover on a scale of 0 to 10, where 0 is open water and 10 is complete ice coverage (fully consolidated). We also identified different types of ice, including first-year, multi-year, and thin gray ice. An automatic logger recorded the ice conditions every 10 s, updated by the observer as conditions changed.

### Hydrology

We measured hydrology at each survey station. Conductivity and temperature in the water column were measured using a Sea-Bird Inc. SBE-9/11+ CTD profiler (Table 1). We used Niskin sampling bottles to collect water samples for measuring chlorophyll *a* (chl *a*) and phaeopigment concentrations (Table 1). For full details, see @dietrich2021winter.

## Statistical analysis

We standardized our data to integrate visual surveys (line transects) with oceanographic sampling stations (discrete points). First, we aggregated all transect observations (relative abundance counts of seabirds and mammals, and ice conditions) into 1.85 km bins. We then associated each sampling station with all transect intervals conducted within 15 km and 3 d of the station to form ‘sites’. We excluded intervals farther in space or time than 15 km and 3 d from a sample station from the community analysis. After associating transect intervals and sample stations in sites, we aggregated the relative predator abundance and ice conditions of each site. We normalized relative predator abundance by sampling effort (ind. km^–1^), then applied a pseudo-log transformation (i.e. log~10~(*x* + 1)) to minimize the influence of highly abundant taxa. The analysis of macrozooplankton in these surveys [@dietrich2021winter] excluded taxa observed at fewer than 5% of sample stations; the seabird and marine mammal assemblage is much less diverse than macrozooplankton, so we did not exclude any species. We aggregated ice conditions across intervals within a site as the most frequent ice type (open, thin, first-year, or multi-year sea ice) and mean percent coverage.

We assessed the community structure from relative abundance data using cluster and ordination analyses. First, we calculated a Bray-Curtis dissimilarity matrix using the R package 'vegan' [@oksanenVeganCommunityEcology2022] to estimate dissimilarities in species composition between sites. We then generated hierarchical clusters using Ward's minimum variance linkage method, implemented in the R package 'cluster' [@maechlerClusterClusterAnalysis2022], applied to the dissimilarity matrix. We determined the optimal number of clusters (*k*) to retain from the hierarchy using the gap statistic [@tibshiraniEstimatingNumberClusters2001]. The gap statistic compares the within-cluster variation of different values of *k* to their expected values under a null reference distribution. The gap statistic, *f*, usually increases with *k*, but exhibits a shoulder at the optimal number of clusters. A recommended heuristic for finding the shoulder is selecting the smallest *k* such that the gap statistic for *k*, *f*(*k*), is greater than the gap statistic for *k + 1* less the standard error, *s* [@eq-gap-stat].

$$
f(k)>f(k+1)-s(k+1)
$$ {#eq-gap-stat}

We described each predator cluster with a smaller number of indicator species. Using the R package 'labdsv', we calculated indicator values, *d*, which indicate the relative affinity of a species to a particular cluster [@dufrene1997species; @robertsLabdsvOrdinationMultivariate2019]. The indicator value for species *i* in cluster *c*, *d~ic~*, is the product of relative frequency, *f~ic~*, and relative abundance, *a~ic~*, where *f~ic~* is the fraction of sites in *c* where *i* was present and *a~ic~* is the ratio of per-site abundance of *i* in *c* to the overall per-site abundance of *i*. *d* ranges between 0 and 1, where values close to 1 indicate a species that was ubiquitous within the cluster and highly abundant relative to other clusters. Indicator values ≥0.25 were retained as significant indicators.

To explore the relationship between community structure and the environment, we reduced the dimensionality of the community data using non-metric multidimensional scaling (NMDS). For this purpose, we used the same Bray-Curtis dissimilarity matrix from the cluster analysis. We determined the number of NMDS axes to retain by examining a scree plot, which plots overall stress of the fit against the number of axes, and assessed quality of fit using a Shepard diagram, which plots the dissimilarity distances in the reduced space against the original dissimilarity distances. We chose the smallest number of axes with stress \<0.2 and examined the correlation in the Shepard diagram for goodness of fit.

We used linear regression to quantify the relationship between the retained NMDS axes and the environmental, biological, and temporal variables. We performed this analysis with the `envfit()` function in the 'vegan' package, which treats environmental variables as the response variables and NMDS ordination scores as the predictors. The environmental variables included physical conditions (hydrology and ice conditions), biology (macrozooplankton community), and temporal variables (Table 1). Temporal variables included time of day of the net tow, to account for diel patterns in macrozooplankton distribution, as well as year, to account for interannual variability. The regression coefficients obtained from the analysis represent the direction and magnitude of the maximal change in community structure associated with each environmental variable. We tested for significant associations between predator and macrozooplankton communities using a chi-squared test and assessed the statistical significance of individual predator–prey community associations via post hoc analysis of the Pearson residuals.

# Results

<!-- Here's some example analysis code: -->

```{r}
#| label: get-data
#| eval: false
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r}
#| label: fig-demo-plot
#| fig-cap: "A plot of random numbers"
plot(rnorm(10))
```

@fig-demo-plot shows how we can have a caption and cross-reference for a plot. Note that figure label and cross-references must both be prefixed with `fig-`

```{r}
#| label: demo-inline-code
#| echo: false
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage