### 2024-06-13 README.Rmd
Update README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Sea-ice and macrozooplankton distribution as determinants of top predator community structure in Antarctic winter"
```

# antwincomm

[![DOI](https://zenodo.org/badge/doi/10.5281/zenodo.11397472.svg)](http://dx.doi.org/10.5281/zenodo.11397472)

This repository contains the code for our paper:

> Czapanskiy MF, Santora JA, Dietrich KS, Cimino MA, Hazen EL, Reiss CS, Veit RR (2024). _`r Title`_. Marine Ecology Progress Series <https://doi.org/10.3354/meps14609>

### How to cite

Please cite this compendium as:

> Czapanskiy MF, Santora JA, Dietrich KS, Cimino MA, Hazen EL, Reiss CS, Veit RR (2024). (`r format(Sys.Date(), "%Y")`). _Compendium of R code for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.5281/zenodo.11397472>

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): Quarto source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. It also has a rendered version,
    `paper.docx`, suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations
  - [:file\_folder: scripts](/analysis/scripts): R scripts for processing data and performing the analysis
  - [:file\_folder: data](/analysis/data): Raw and derived data

## How to download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip from from this URL:
[main.zip](/archive/main.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
[DESCRIPTION](/DESCRIPTION) file).
- download the raw data (see [/analysis/data/raw_data/README.md](/analysis/data/raw_data/README.md))
- run all the data processing and analysis scripts (run [/analysis/scripts/_make_all.R](/analysis/scripts/_make_all.R))
- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file